visitation_bias

### 2021-04-28 Dryad_ms1/README.Rmd
switched to proportional visitation data for null models
resurrected lability analysis

@@

>>>>

## 4. `bb_traits.csv`

### Description: Bumble bee trait data

Clade and subgenus classifications are based on Williams et al (2008).

Tongue length data are from Arbetman et al. (2017), Obeso (1992), and Durieux (2000)

### Columns:
 1. `clade` = bumble bee clade: LF = long faced, SH = short faced, M = Mendacibombus
 2. `subgenus` = bumble bee subgenus
 3. `bb.sp.lat` = full bumble bee species name
 4. `bb.sp` = abbreviated bumble bee species name
 5. `pbl.w` = worker tongue length based primarily on Arbetman et al. (2017) and supplemented where necessary from Durieux (2000) and Obeso (1992)
 6. `pbl.w.class` = discretized worker tongue length
 7. `pbl.w.class2` = simplified discretized worker tongue length
 8. `pbl.w.ref` = reference from which pbl.w was gathered
 9. `Notes` = notes on pbl.w
 10. `pbl.w.durieux` = alternative set of tongue length based primarily on Durieux (2000) and supplemented where necessary from Arbetman et al. (2017) and Obeso (1992)
 11. `notes_durieux` = notes on pbl.w.durieux

## 5. `fl_traits.csv`

### Description: Kugler's (1970) floral morphotype classification for floral species

For a description of Kugler's morphotype codes, see `kugler_key.tsv`.

Data were accessed in April 2020 via the Bioflor plant trait database (Klotz et al. 2002). Missing data were manually supplemented as descrtibed in `fl_trait_proc.R`.

### Columns:
 1. `plant.sp` = plant species
 2. `plant.genus` = plant genus
 3. `plant.family` = plant family
 4. `k.type` = Kugler morphotype
 5. `k.type.s` = Kugler morphotype simplified to 1 decimal place
 6. `k.type.ss` = Kugler morphotype simplified to 0 decimal places

## 6. `kugler_key.tsv`

### Description: A verbal description of Kugler's (1970) floral morphotype classification

Data were accessed in April 2020 via the Bioflor plant trait database (Klotz et al. 2002).

### Columns:
 1. `code` = Morphotype code
 2. `description` = Morphotype description

@@

## 2. `fl_trait_proc.R`

### Description: R script for processing raw BIOFLOR trait data and adding traits for missing species.

# References

Williams PH, Cameron SA, Hines HM, Cederberg B, Rasmont P. 2018. A simplified subgeneric classification of the bumblebees (genus Bombus). Apidologie, 39, pp. 46-74

Arbetman MP, Gleiser G,Morales CL, Williams P, Aizen MA. 2017 Global decline of bumblebees is phylogenetically structured and inversely related to species range size and pathogen incidence. Proc. R. Soc. B 284: 20170204.

Obeso JR. 1992. Geographic distributions and community structure of Bumblebees in Northern Iberian Peninsula. Oecologia, 89, 244–252.

Durieux E-A. 2000. Etude des choix floraux des bourdons (Hymenoptera, Apidae) de la commune d’eyne (france, Pyrénées-Orientales). PhD thesis, Université de Mons-Hainaut.

Kugler, H. 1970. Blütenökologie. Gustav Fischer, Stuttgart.

Klotz, S, Kühn I, Durka W. 2002. BIOLFLOR - eine datenbank zu biologisch-ökologischen merkmalen der gefäßpflanzen in deutschland. https://www.ufz.de/index.php?en=38567

### 2021-04-28 code/ms2_ktype_preference_nullmodel_analysis.Rmd
resurrected lability analysis

>>>>

# 3. Null model 3: bb.sp x ktype, by site-date
```{r}
# Nest and join consumer and resource data frames

# For HPC
library(tidyverse)
library(econullnetr)
prop+visits_ktype <- read_csv("./prop_visits_ktype.csv")

con_nest2 <- prop_visits_ktype %>%
  select(site, date, bb.sp, ktype, prop.visits) %>%
  pivot_wider(values_from = prop.visits, names_from = ktype, values_fill = 0) %>%
  select(-c("0", "10", "9")) %>%
  group_by(site, date) %>%
  nest() %>%
  rename(bees = data)

res_nest2 <- prop_visits_ktype %>%
  select(site, date, ktype, perc.flower.cover) %>%
  distinct() %>%
  pivot_wider(values_from = perc.flower.cover, names_from = ktype, values_fill = 0) %>%
  select(-c("0", "10", "9")) %>%
  group_by(site, date) %>%
  nest() %>%
  rename(plants = data)

conres2 <- full_join(con_nest2, res_nest2)

# Create a big beautiful table of data and models (run on HPC!)
# This will probably take 8 hours to run
null3 <- conres2 %>%

  # Generate null models
  mutate(null.model = map2(bees, plants, function(x, y)
    generate_null_net(consumers = x[, 1:8],
                      resources = y[, 1:7],
                      sims = 1000, # raised from default of 100
                      data.type = "quantities",
                      prog.count = TRUE)))
write_rds(null3, "./null3.rds") # hpc
null3 <- readRDS("../output/null3.rds")

# Tabulate results (with test)
null_tab3 <- null3 %>%
  mutate(null.model.tab = map(null.model, function(x)
    test_interactions(x, signif.level = 0.99) %>%
      as_tibble() %>%
      mutate(preference = Observed - Null,
             preference.upper99 = Observed - Upper.99.CL,
             preference.lower99 = Observed - Lower.99.CL) %>%
      select(bb.sp = Consumer, ktype = Resource, everything()))) %>%

  # Unnest and add site data and bumble bee trait data
  unnest(null.model.tab) %>%
  select(site, date, bb.sp, ktype, Observed, Null, Lower.99.CL, Upper.99.CL, Test, 
         SES, preference, preference.upper99, preference.lower99) %>%
  left_join(site_data) %>%
  left_join(bb_traits) %>%
  group_by(site, date, bb.sp) %>%
  mutate(total.observed = sum(Observed)) %>%
  filter(total.observed > 0) %>% # for each species, drop sites where the species did not occur
  filter(Observed != 0 | Null != 0) %>% # when both equal zero, it's probably because the ktype had zero abundance; in any case, it's meaningless to retain these, and they just cause zero-inflation
  ungroup() %>%
  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
  mutate(ktype = factor(ktype),
         site = factor(site),
         yday = yday(date),
         year = factor(year(date)),
         bb.ktype = factor(bb.ktype),
         bb.sp = factor(bb.sp, levels = c("jone", "pyre", "soro", "psit", # order by ascending tongue length
                                          "telu", "lapi", "prat", "mont",
                                          "wurf", "muci", "pasc", "hypn",
                                          "mend", "hort", "gers")))

gam3 <- bam(preference ~
              bb.ktype +
              s(site, by = bb.ktype, bs = "re") +
              s(year, by = bb.ktype, bs = "re") +
              te(yday, elev.mean, by = bb.ktype),
            family = "gaussian",
            method = "fREML",
            discrete = TRUE,
            data = filter(null_tab3, 
                          bb.sp %in% c("pasc", "prat", "telu", 
                                       "wurf", "soro", "hort") &
                            ktype %in% c("3", "5", "6", "7"))) %>% 
              getViz()

check.gamViz(gam3)
gam3_sum <- summary(gam3)
plot.gamViz(gam3, allTerms = TRUE)

hort_3 <- plot(sm(gam3, 49)) + 
  labs(y = NULL, x = NULL) + 
  theme(legend.position = "top") + 
  theme_void(20)

hort_5 <- plot(sm(gam3, 50)) + labs(y = NULL, x = NULL) + theme_void(20)
hort_6 <- plot(sm(gam3, 51)) + labs(y = NULL, x = NULL) + theme_void(20)
hort_7 <- plot(sm(gam3, 52)) + labs(y = NULL, x = NULL) + theme_void(20)

pasc_3 <- plot(sm(gam3, 53)) + labs(y = NULL, x = NULL) + theme_void(20)
pasc_5 <- plot(sm(gam3, 54)) + labs(y = NULL, x = NULL) + theme_void(20)
pasc_6 <- plot(sm(gam3, 55)) + labs(y = NULL, x = NULL) + theme_void(20)
pasc_7 <- plot(sm(gam3, 56)) + labs(y = NULL, x = NULL) + theme_void(20)

prat_3 <- plot(sm(gam3, 57)) + labs(y = NULL, x = NULL) + theme_void(20)
prat_5 <- plot(sm(gam3, 58)) + labs(y = NULL, x = NULL) + theme_void(20)
prat_6 <- plot(sm(gam3, 59)) + labs(y = NULL, x = NULL) + theme_void(20)
prat_7 <- plot(sm(gam3, 60)) + labs(y = NULL, x = NULL) + theme_void(20)

soro_3 <- plot(sm(gam3, 61)) + labs(y = NULL, x = NULL) + theme_void(20)
soro_5 <- plot(sm(gam3, 62)) + labs(y = NULL, x = NULL) + theme_void(20)
soro_6 <- plot(sm(gam3, 63)) + labs(y = NULL, x = NULL) + theme_void(20)
soro_7 <- plot(sm(gam3, 64)) + labs(y = NULL, x = NULL) + theme_void(20)

telu_3 <- plot(sm(gam3, 65)) + labs(y = NULL, x = NULL) + theme_void(20)
telu_5 <- plot(sm(gam3, 66)) + labs(y = NULL, x = NULL) + theme_void(20)
telu_6 <- plot(sm(gam3, 67)) + labs(y = NULL, x = NULL) + theme_void(20)
telu_7 <- plot(sm(gam3, 68)) + labs(y = NULL, x = NULL) + theme_void(20)

wurf_3 <- plot(sm(gam3, 69)) + labs(y = NULL, x = NULL) + theme_void(20)
wurf_5 <- plot(sm(gam3, 70)) + labs(y = NULL, x = NULL) + theme_void(20)
wurf_6 <- plot(sm(gam3, 71)) + labs(y = NULL, x = NULL) + theme_void(20)
wurf_7 <- plot(sm(gam3, 72)) + labs(y = NULL, x = NULL) + theme_void(20)

pdf("../output/gam3_figures/hort_3.pdf") 
hort_3 
dev.off()

pdf("../output/gam3_figures/hort_5.pdf") 
hort_5 
dev.off()

pdf("../output/gam3_figures/hort_6.pdf") 
hort_6 
dev.off()

pdf("../output/gam3_figures/hort_7.pdf") 
hort_7 
dev.off()

pdf("../output/gam3_figures/pasc_3.pdf") 
pasc_3 
dev.off()

pdf("../output/gam3_figures/pasc_5.pdf") 
pasc_5 
dev.off()

pdf("../output/gam3_figures/pasc_6.pdf") 
pasc_6 
dev.off()

pdf("../output/gam3_figures/pasc_7.pdf") 
pasc_7 
dev.off()

pdf("../output/gam3_figures/prat_3.pdf") 
prat_3 
dev.off()

pdf("../output/gam3_figures/prat_5.pdf") 
prat_5 
dev.off()

pdf("../output/gam3_figures/prat_6.pdf") 
prat_6 
dev.off()

pdf("../output/gam3_figures/prat_7.pdf") 
prat_7 
dev.off()

pdf("../output/gam3_figures/soro_3.pdf") 
soro_3 
dev.off()

pdf("../output/gam3_figures/soro_5.pdf") 
soro_5 
dev.off()

pdf("../output/gam3_figures/soro_6.pdf") 
soro_6 
dev.off()

pdf("../output/gam3_figures/soro_7.pdf") 
soro_7 
dev.off()

pdf("../output/gam3_figures/telu_3.pdf") 
telu_3 
dev.off()

pdf("../output/gam3_figures/telu_5.pdf") 
telu_5 
dev.off()

pdf("../output/gam3_figures/telu_6.pdf") 
telu_6 
dev.off()

pdf("../output/gam3_figures/telu_7.pdf") 
telu_7 
dev.off()

pdf("../output/gam3_figures/wurf_3.pdf") 
wurf_3 
dev.off()

pdf("../output/gam3_figures/wurf_5.pdf") 
wurf_5 
dev.off()

pdf("../output/gam3_figures/wurf_6.pdf") 
wurf_6 
dev.off()

pdf("../output/gam3_figures/wurf_7.pdf") 
wurf_7 
dev.off()
```
